% !TEX root = ../mainthesis.tex

%Acknowledgments

\renewcommand{\baselinestretch}{2}
\small\normalsize
\hbox{\ }

\vspace{-.65in}

\begin{center}
\large{Acknowledgments}
\end{center}

\vspace{1ex}

I would like to start off by thanking my graduate advisor, Dr. Eric Wachsman, for his support during my graduate career.
He allowed me the freedom to direct the details of my research and the support to help me reach my goals.
He was also supportive of  me when my family began to grow through marriage and children, allowing me to continue to develop my personal life while pursuing my degree.

More thanks than I can express is due to my wonderful wife, Bethany and now our daughters, Lucy and Madeleine.
When Bethany and I first met, I was still deciding between gradate programs, and now, after many life changes, she continues to challenge and support me in my endeavors.
Lucy and Madeleine have provided me with impetus to continue pushing through the challenges and to strike a balance between my personal and professional pursuits.

Of course I owe acknowledgement to my parents, Jim and Marilyn Stanley, who raised me to be inquisitive and persistent.
Thank you for helping me to become the person that I am.

The Materials Science and Engineering Department here at UMD has been filled with marvelous and helpful people.
Dr. Kathleen Hart was one of the first people I met in the department, seeming to help organize everything, always being there when needed.
She has been sorely missed this past year and I hope she is making the most of her retirement.
Kay Morris and Jenna Bishop in the business office help the department run smoothly, if it be answering benefits and procurement questions, or doting on my children when they visited.
Then of the many professors I have been able to interact with, I would like to especially thank Dr. Isabel Lloyd who has graciously given me her advice, help make introductions, and passed along opportunities to me when they come her way.

Somehow a small group of my friends of undergraduate have accumulated in the area.
It has been a blessing to have them near by, always willing to help out when need be or to just spend time with each other.
Dave and Katie Shahin are a large part of what brought me to the University and it has been a blast to go through the program with Dave.
Andrew Dunkman has helped me in so many ways, from apartment hunting a thousand miles away to programming questions, for which I express my thanks.
And finally, Angela Rudolph has been very supportive of me and my family getting through graduate school.

I owe a good amount of thanks to the various people in lab who helped show me the way to get started with my experiments, allowed me to bounce ideas off them, or helped with the work.
Chris Pellegrinelli and Greg Hitz, you were two people who I was always comfortable with going to and asking what seemed like the most basic questions on how to do things, helping me get started in lab.
Mohammad Hussain and Yi-Lin Huang shared their experience and knowledge of material science as I worked through the fundamentals.
Thanks to Tom Hays for being my partner as we started mechanical testing in lab.
Ian Robinson, Evans Gritton and Jon O'Neill provided much needed help with SEM, XRD, and modeling, saving me from fumbling through it myself.

I would like to acknowledge my friends and classmates at the University who formed part of my support network, sharing knowledge and encouragement.
Mimi Hiebert, Doug Henderson, Michael Van Order, Beth Tennyson, Travis Dietz, Adam Pranda, Marina Pranda, and Max Lerman, thank you for the time spent together in and out of class as we worked through our problems together laughing and having fun along the way.

Early when starting to look into mechanical properties, I was introduced to Mr. George Quinn at the National Institute of Standards and Technology.
Mr. Quinn was kind enough to take much of his time to look though my preliminary work and give me feedback on testing methodology and fractography.
It was a great privilege to meet the man who literally wrote the standards which I was attempting to follow and to have his guidance as I started to navigate my way though his world.

Thank you to Dr. Sean Bishop, who helped guide me as I made my ambitious goals and helped me plan out the steps needed to accomplish them.

Many facilities across campus assisted in the collection and analysis of data for this work. My thanks go to the Advanced Imaging \& Microscopy Laboratory (AIMLab), the X-Ray Crystallographic Center, the Surface Analysis Center, Dr. Robert Bonenberger at the Modern Engineering Materials Instructional Laboratory,

Thank you to Redox Power Systems, the Department of Energy's National Energy Technology Laboratory and the Maryland Industrial Partnerships Program for supporting this work.
